---
title: "Forest Fires in Brazil"
author: "Carlos Habekost"
date: "18/11/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r message=FALSE, warning=FALSE, echo=FALSE}

#Packages required
if(!require(ggplot2)){install.packages("ggplot2")}
if(!require(hrbrthemes)){install.packages("hrbrthemes")}
if(!require(plotly)){install.packages("plotly")}

if(!require(tidyverse)){install.packages("tidyverse")}
if(!require(readr)){install.packages("readr")}

options(scipen = 10000)

```

## Introduction

This document is aiming to analysis the forest fires that occurred in Brazil from 1998 to 2017. Forest Fires are a critical problem for the Tropical Forest in Brazil. I obtained this data from [kaggle.com](https://www.kaggle.com/gustavomodelli/forest-fires-in-brazil), who organized the data initially derived from the official website of the [Brazilian government](http://dados.gov.br/dataset/sistema-nacional-de-informacoes-florestais-snif).

With this analysis, we can identify fires occurred in Brazilian states, considering: Year, month e number of Fires happened.

## Analisys Performed

The original data set has 6,444 rows. The following columns were identified:

- **year:** when Forest Fires happened
- **state:** Brazilian State
- **month:** Month when Forest Fires happened
- **number:** Number of Forest Fires reported
- **date:** Date when Forest Fires where reported

I had a pre-processing in the data set, aiming to perform correctly in this markdown:

- *Months*: Originally was in Portuguese, I standardization in English.
- *Standardization of State name accents*: Originally in this data set, the name of some States had an accent; I normalization without accent to avoid issues during the execution of this markdown.

```{r}
reportFires <- read.csv(file="amazon.csv")

#print(reportFires)
```

With the data ready, I first summarize some information about the dataset:

```{r}
summary(reportFires)
```

Second, using ggplot, I plot a heatmap aiming to identify the most reported state with fires. It is evident that the state of Mato Grosso is most affected by fires (totaling 250000+ fires reported):

```{r}
library(dplyr)

analyzedData2 <- reportFires %>% select(-date) %>%
  group_by(state, month) %>%
  summarise(totalFires = sum(number))


var <- analyzedData2 %>%
  mutate(month = factor(month, levels = c("January", "February","March", "April", "May", "June", "July", "August", "September", "October", "November", "December"))) 
  #%>%
  
ggplot(var, aes(x=month,y=reorder(state,desc(state)),fill=totalFires)) + 
  geom_tile() +
  scale_fill_gradient(low = "white", high = "red") +
  theme(axis.text.x = element_text(angle = 90, vjust=0.25)) +
  xlab( "Month" ) +
  ylab( "States" ) +
  labs(fill="Total Fires")

```

In the third analysis, I investigate the fire reported by the state. As a result, considering the period analyzed, Mato Grosso has the most number of fires (with 767,890), Para (577,481) and Maranhão (375,279).

```{r}
# Library
library(ggplot2)

var %>% group_by(state) %>% summarise(totalFires = sum(totalFires)) %>%
ggplot(aes(x=state,y = totalFires, fill=totalFires)) +
  geom_col() +
  xlab( "State" ) +
  ylab( "Number of Fires" )+
  scale_fill_gradient(low = "green", high = "red") +
  geom_text(aes(label=totalFires), vjust=-0.3) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, vjust=0.25), legend.position = "none")

```


In the fourth analysis, I considered the month with more cases of fire, considering all fires reported. As a result, I identified that September has the most number of fire reports (totalizing 1,015,925), after August (740,841 reports) and October (629,665 reports).


```{r}
# Library
library(ggplot2)

var %>% group_by(month) %>% summarise(totalFires = sum(totalFires)) %>%
ggplot(aes(x=month,y = totalFires, fill=totalFires)) +
  geom_col() +
  xlab( "Month" ) +
  ylab( "Number of Fires" )+
  scale_fill_gradient(low = "green", high = "red") +
  geom_text(aes(label=totalFires), vjust=-0.3) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, vjust=0.25), legend.position = "none")

```

Finally, I tried to perform other analyzes using ggplot. However, I did not obtain an expressive result to include in this work. 

As an example, I present an analysis considering the state of the Mato Grosso, using a histogram graph. However, the obtained graph has a few analysis values. Also, I tried to plot this graph, considering the data for all states. However, the results obtained were similar to the heatmap graph.

```{r}
# Library
library(ggplot2)

reportFires %>% filter(state =="Mato Grosso") %>%
  ggplot(aes(x=number)) +
  geom_histogram(fill = 'red', bins=30) +
  theme_bw()
```




